---
title: "FinalProject_PreProcessing"
author: "Richa"
date: "2024-04-22"
output: html_document
---
# Project Objective: 

The overarching objective of this project is to develop a model to predict hospital readmission based on key characteristics from the patient’s first visit. 
Our goal is to compare several statistical learning methods and their test performance and determine which model will most accurately predict whether or not someone will be readmitted at a later time (within a 2-year time frame, given the structure of the data).

#### load libraries

```{r, warning=FALSE, message=FALSE}
library(tidyr)
library(tidyverse)
library(dplyr)
library(VIM)
library(lubridate)
library(readr)
```

#### read in data

```{r}
admission <- read.csv("data/HDHI Admission data.csv")
mortality <- read.csv("data/HDHI Mortality Data.csv")
```

```{r, warning = FALSE}
# convert columns to numeric
admission <- admission %>% 
  filter(MRD.No. != "NILL") %>% 
  mutate(across(c(MRD.No., HB:BNP, EF, CHEST.INFECTION), ~as.numeric(.))) 
```

#### knn with k = 10 for missing predictors

```{r}
admission_columns <- colnames(admission)
```

```{r}
# replacing blanks with NAs to be able to impute using KNNs
admission <- replace(admission, admission == "", NA)
admission <- replace(admission, admission == "EMPTY", NA)
admission <- replace(admission, admission == "\\", NA)
sum(is.na(admission))
```


```{r}
admission_imp <- kNN(admission, k = 5, variable = admission_columns, imp_var = FALSE)
```

```{r}
sum(is.na(admission_imp))
```

#### length of stay - 3 options:

- use original continous variable
- convert to categorical quantiles
- cut continuous varibale at 15

##### categorical quantiles

```{r}
# cargorical quantiles
cutoffs = quantile(admission_imp$DURATION.OF.STAY, probs = seq(0, 1, 0.25))
admission_imp <- admission_imp %>% 
  mutate(Duration_Cat = cut(DURATION.OF.STAY, breaks = cutoffs)) 
```

#### generate the outcome variable column: 

- limiting to only those with readmission and only their first visit

```{r}
processed_admissions <- admission_imp %>% 
  #mutate(D.O.A = coalesce(as.Date(D.O.A, format = "%m/%d/%Y")), as.Date(D.O.A, format = "%d/%m/%Y")) %>% 
  mutate(D.O.A = coalesce(mdy(D.O.A), dmy(D.O.A))) %>% 
  arrange(MRD.No., D.O.A) %>% 
  group_by(MRD.No.) %>% 
  mutate(num_visits = n()) %>% 
  slice(1) %>% 
  ungroup() %>% 
  filter(num_visits > 1)
```

```{r}
sum(is.na(processed_admissions$D.O.A))
```

#### train test split 80/20

```{r}
set.seed(263)
indices <- sample(1:nrow(processed_admissions), size = floor(0.8 * nrow(processed_admissions)), replace = FALSE)
admissions_train <- processed_admissions[indices, ]
admissions_test <- processed_admissions[-indices, ]
```

```{r}
saveRDS(admissions_train, "admissions_train.rdata")
saveRDS(admissions_test, "admissions_test.rdata")
```


#### models: 
- forest related 
- bagging
- penalized regression 
- fit logistic to predict mortality and use those probabilities as weights for IPW
- GAMs 






