---
title: "FinalProject_PreProcessing"
author: "Richa"
date: "2024-04-22"
output: html_document
---
# Project Objective: 

The overarching objective of this project is to develop a model to predict hospital readmission based on key characteristics from the patient’s first visit. 
Our goal is to compare several statistical learning methods and their test performance and determine which model will most accurately predict whether or not someone will be readmitted at a later time (within a 2-year time frame, given the structure of the data).

#### load libraries

```{r, warning=FALSE, message=FALSE}
library(tidyr)
library(tidyverse)
library(dplyr)
library(VIM)
library(lubridate)
library(readr)
```

#### read in data

```{r}
admission <- read.csv("data/HDHI Admission data.csv")
mortality <- read.csv("data/HDHI Mortality Data.csv")
```

```{r, warning = FALSE}
# convert columns to numeric
admission <- admission %>% 
  filter(MRD.No. != "NILL") %>% 
  mutate(across(c(MRD.No., HB:BNP, EF, CHEST.INFECTION), ~as.numeric(.))) 
```

#### knn with k = 10 for missing predictors

```{r}
admission_columns <- colnames(admission)
```

```{r}
# replacing blanks with NAs to be able to impute using KNNs
admission <- replace(admission, admission == "", NA)
admission <- replace(admission, admission == "EMPTY", NA)
admission <- replace(admission, admission == "\\", NA)
sum(is.na(admission))
```

```{r}
admission_imp <- kNN(admission, k = 5, variable = admission_columns, imp_var = FALSE)
```

```{r}
sum(is.na(admission_imp))
```

<!-- #### length of stay - 3 options: -->

<!-- - use original continuous variable -->
<!-- - convert to categorical quantiles -->
<!-- - cut continuous variable at 15 -->

<!-- ##### categorical quantiles -->

<!-- # Note: Individuals with duration stay = 1 are categorized as NA bc buckets are not left inclusive -->

<!-- ```{r} -->
<!-- # categorical quantiles -->
<!-- cutoffs = quantile(admission_imp$DURATION.OF.STAY, probs = seq(0, 1, 0.25)) -->
<!-- admission_imp <- admission_imp %>%  -->
<!--   mutate(Duration_Cat = cut(DURATION.OF.STAY, breaks = cutoffs))  -->
<!-- ``` -->

#### generate the outcome variable column: 

- limiting to only those with readmission and only their first visit

```{r}
processed_admissions <- admission_imp %>% 
  mutate(D.O.A = coalesce(mdy(D.O.A), dmy(D.O.A))) %>% 
  arrange(MRD.No., D.O.A) %>% 
  group_by(MRD.No.) %>% 
  mutate(num_visits = n()) %>% 
  dplyr::slice(1) %>% 
  ungroup() %>% 
  mutate(readmission = ifelse(num_visits > 1, 1, 0))
```

```{r}
sum(is.na(processed_admissions$D.O.A))
```

#### train test split 80/20

```{r}
set.seed(263)
indices <- sample(1:nrow(processed_admissions), size = floor(0.8 * nrow(processed_admissions)), replace = FALSE)
admissions_train <- processed_admissions[indices, ]
admissions_test <- processed_admissions[-indices, ]
```

```{r}
saveRDS(processed_admissions, "all_admissions.rdata")
saveRDS(admissions_train, "admissions_train.rdata")
saveRDS(admissions_test, "admissions_test.rdata")
```





